### 2024-05-25 getting-started.qmd
20240525 - calculate player age

```{r}
library("nflreadr")
library("nflfastR")
library("progressr")
library("tidyverse")
```

@@

@@

load(file = here("data", "nfl_expectedFantasyPoints_weekly.RData"))

load(file = here("data", "nfl_expectedFantasyPoints_offense.RData"))
```

@@

>>>>

```{r}
library("nflreadr")
library("nflfastR")
library("progressr")
library("lubridate")
library("tidyverse")
```

@@

load(file = here("data", "nfl_players.RData"))
load(file = here("data", "nfl_teams.RData"))
load(file = here("data", "nfl_rosters.RData"))
load(file = here("data", "nfl_rosters_weekly.RData"))
load(file = here("data", "nfl_schedules.RData"))
load(file = here("data", "nfl_combine.RData"))
load(file = here("data", "nfl_draftPicks.RData"))
load(file = here("data", "nfl_depthCharts.RData"))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
load(file = here("data", "nfl_actualStats_weekly.RData"))
load(file = here("data", "nfl_injuries.RData"))
load(file = here("data", "nfl_snapCounts.RData"))
load(file = here("data", "nfl_espnQBR.RData"))
load(file = here("data", "nfl_nextGenStats_weekly.RData"))
load(file = here("data", "nfl_advancedStatsPFR.RData"))
load(file = here("data", "nfl_playerContracts.RData"))
load(file = here("data", "nfl_ftnCharting.RData"))
load(file = here("data", "nfl_playerIDs.RData"))
load(file = here("data", "nfl_rankings.RData"))
load(file = here("data", "nfl_expectedFantasyPoints_weekly.RData"))
load(file = here("data", "nfl_expectedFantasyPoints_offense.RData"))
```

@@

load(file = here("data", "nfl_expectedFantasyPoints_weekly.RData"))
load(file = here("data", "nfl_expectedFantasyPoints_offense.RData"))
```

@@

### Player Age {#sec-calculateActualPoints}

```{r}
# Reshape from wide to long format
nfl_actualStats_offense_weekly_long <- nfl_actualStats_offense_weekly %>% 
  pivot_longer(
    cols = c(recent_team, opponent_team),
    names_to = "role",
    values_to = "team")

# Reshape from wide to long format
nfl_schedules_long <- nfl_schedules %>% 
  pivot_longer(
    cols = c(home_team, away_team),
    names_to = "role",
    values_to = "team")

# Perform separate inner join operations for the home_team and away_team
nfl_actualStats_offense_weekly_home <- inner_join(
  nfl_actualStats_offense_weekly_long,
  nfl_schedules,
  by = c("season","week","team" = "home_team")) %>% 
  mutate(home_away = "home_team")

nfl_actualStats_offense_weekly_away <- inner_join(
  nfl_actualStats_offense_weekly_long,
  nfl_schedules,
  by = c("season","week","team" = "away_team")) %>% 
  mutate(home_away = "away_team")

# Combine the results of the join operations
nfl_actualStats_offense_weekly_schedules_long <- bind_rows(
  nfl_actualStats_offense_weekly_home,
  nfl_actualStats_offense_weekly_away)

# Reshape from long to wide
nfl_actualStats_offense_weekly_schedules <- nfl_actualStats_offense_weekly_schedules_long %>%
  distinct(player_id, season, week, game_id, home_away, team, gameday) %>% #, .keep_all = TRUE
  pivot_wider(
    names_from = home_away,
    values_from = team)

# Merge player birthdate and the game date
player_birthdate_gameday <- left_join(
  nfl_actualStats_offense_weekly_schedules,
  unique(nfl_players[,c("gsis_id","birth_date")]),
  by = c("player_id" = "gsis_id")
)

player_birthdate_gameday$birth_date <- ymd(player_birthdate_gameday$birth_date)
player_birthdate_gameday$gameday <- ymd(player_birthdate_gameday$gameday)

# Calculate player's age for a given week as the difference between their birthdate and the game date
player_birthdate_gameday$age <- interval(
  start = player_birthdate_gameday$birth_date,
  end = player_birthdate_gameday$gameday
) %>% 
  time_length(unit = "years")

# Merge with player info
player_age <-  left_join(
  player_birthdate_gameday,
  nfl_players %>% select(-birth_date),
  by = c("season", "player_id" = "gsis_id"))

# Add game_id to weekly stats to facilitate merging
nfl_actualStats_offense_weekly <- nfl_actualStats_offense_weekly %>% 
  left_join(
    player_age[,c("season","week","player_id","game_id")],
    by = c("season","week","player_id"))

# Merge with player weekly stats
player_age_stats <- left_join(
  player_age,
  nfl_actualStats_offense_weekly,
  by = c(c("season","week","player_id","game_id")))

player_age_stats$years_of_experience <- as.integer(player_age_stats$years_of_experience)
```
